Lab3 P2
```{r}
install.packages("sqldf")
```

package used
```{r}
require(stringr)
require(sqldf)
```

1. Data preparation and omit the row name.
```{r}
inputTable <- read.csv(file.choose())
inputTable <- inputTable[,-1]
```

2. Unified the value (Data Cleasing)
```{r}
clean <- inputTable
#Seperate Value and Unit
clean$BOD <- as.Date(clean$BOD, format = '%d/%m/%Y') # split the date 

# Seperate Value and Unit
clean$Hvalue <- as.numeric(str_extract(clean$Height,"\\d+\\.?\\d*")) # split Height col 
clean$Hunit <- str_extract(clean$Height,"[:alpha:]+") # split Height col 
clean$Height <- NULL # make it null no val

clean$Wvalue <- as.numeric(str_extract(clean$Weight,"\\d+\\.?\\d*")) # split Weight col 
clean$Wunit <- str_extract(clean$Weight,"[:alpha:]+") # split Weight col 
clean$Weight <- NULL

#transform value under unified unit
clean$Height <- unlist(mapply(function(x,y){
  if (y == "m") {
    x <- x*100 # if y is m than x * 100
  }else if (y == "mm") {
    x <- x/10 # if y is mm than x / 10
  }else{
    x <- x # if y is not m or mm than keep it
  }
},clean$Hvalue,clean$Hunit))


clean$Weight <- unlist(mapply(function(x,y){
  if (y == "lb") {
    x <- x/2.2 # if y is lb than x / 2.2
  }else{
    x <- x # if y is not lb than keep it
  }
},clean$Wvalue,clean$Wunit))


clean <- clean[,-c(5:8)]
clean # call the table
```

3. transform value (use sqldf may be better) (ETL)
```{r}
transform <- clean
#transform$Race <- sapply(transform$Race, function(x){
#  if (x==1){
#    x <- "American Indian"
#  }else if (x==2){
#    x <- "Asian"
#  }else if (x==3){
#    x <- "Black"
#  }else if (x==4){
#    x <- "Hispanic"
#  }else if (x==5){
#    x <- "White"
#  }else if (x==6){
#    x <- "Middle-Easten"
#  }
#})

Reference <- data.frame(Code = 1:6,
                        RaceName = c("American Indian",
                                     "Asian",
                                     "Black",
                                     "Hispanic",
                                     "White",
                                     "Middle-Easten"))

# Reference

transform <- merge(transform,
                   Reference,
                   by.x = "Race",
                   by.y = "Code",
                   sort=F)

transform <- transform[,-1]

transform
```

4. Cal Age and BMI (ETL)
```{r}
Final <- transform

Final$BMI <- Final$Weight/(Final$Height/100)^2

Final$Age <- as.numeric(format(Sys.Date(),"%Y"))-as.numeric(format(Final$BOD,"%Y"))

Final
```

5. Average age by gender and show as dataframe.
```{r}
data.frame(AverageAge = tapply(Final$Age, Final$gender,mean))
```

6. Average BMI by Race who are older than 29
```{r}
older29 <- Final[Final$Age>29,]
data.frame(AverageBMI = tapply(older29$BMI,older29$RaceName,mean))
```

7. Find all young adult girls (age around 18~22)
```{r}
query <- Final$gender=="Female"&Final$Age>=27&Final$Age<=29
Final[query,]
```

8. count number of people by gender
```{r}
# tapply(Final$gender, Final$gender, length)

table(Final$gender)
```

9. List all people born in Year 2000
```{r}
Final[format(Final$BOD,"%Y")==2000,]
```

10. Find the details of listed people
```{r}
Namelist <- c("Foster, Jade", "Manegdeg, An", "Knox-Bell, Ahmawn", "Wright, Cheyenne", "al-Beshara, Sajaa", "Watkins, Nykea")

Final[Final$Name %in% Namelist,]
```

11. Different between One of the Tallest Asian and One of the shortest American, And who?
```{r}
Asian <- Final[Final$RaceName=="Asian",]
print(TA <- Asian[Asian$Height==max(Asian$Height),][1,])

# shortest American
USA <- Final[Final$RaceName=="American Indian",]
print(SU <- USA[USA$Height==min(USA$Height),][1,])

print(paste("Different is", TA$Height-SU$Height))
```

12.There are some steps
  1. Find all Asian who are the age from 25 to 30.
  2. list the detail(Name,HValue,WValue) of the targeted people.
  3. sort the pervious result(Name,HValue) by WValue in decending.
```{r}
Result1 <- Final[Final$RaceName=="Asian"&Final$Age>=25&Final$Age<=30,]
Result2 <- Result1[,c("Name","HValue","WValue"),]
# Result3 <- Result2[order(Result2)]
```

The following part using SQL (you may use more than one Query if necessary)

13. Find all mans, age around 24 to 29, and calculate the average Weight by Race.
```{r}
sqldf("
      SELECT RaceName, Avg(Weight) as MeanWeight
      FROM Final
      WHERE gender = 'Male' AND Age Between 24 AND 29
      GROUP BY RaceName
      ")
```

14. List all the Names that their Race average Height(HValue) that higher than 150cm.
```{r}
partial <- sqldf("
      SELECT RaceName, Avg(Height) as MH 
      FROM Final 
      GROUP BY RaceName 
      HAVING MH > 150
      ")

sqldf("
      SELECT Final.Name, Final.RaceName
      FROM Final, partial
      WHERE Final.RaceName = partial.RaceName
      ")
```

15. Calculate average age of this dataset.
```{r}
sqldf("
      SELECT Avg(Age) as MeanAge
      FROM Final
      ")
```

OK! Let try some plottings! This is just a simple trial, more information and learning about Data Visualization will be shown in ITP4864.

16.	Visualise the Race distribution who was born in 1997. 
```{r}
Born1997 <- Final[format(Final$BOD, "%Y")==1997,]
RaceCount <- data.frame(table(Born1997$RaceName))
pie(RaceCount$Freq,
    labels = RaceCount$Var1,
    col=topo.colors(nrow(RaceCount)))
```

